---
title: "Data Science in a Box Lab 5"
author: "Alex Lindquist, Nate Smullen, Wynter Malone, Yadhira Garcia"
date: "10/17/2022"
output: html_notebook
---

##Load Packages

```{r}
library(tidyverse)
library(dsbox)
library(here)
library(janitor)
```

##Alaska
```{r}
dn_ak <- dennys %>%
  filter(state == "AK")
nrow(dn_ak)
```
There are 3 Denny's in Alaska.

```{r}
lq_ak <- laquinta %>%
  filter(state == "AK")
nrow(lq_ak)
```
There are 2 LaQuintas in Alaska.

```{r}
dn_lq_ak <- full_join(dn_ak, lq_ak, by = "state")
dn_lq_ak
```
There are 6 observations in this joined dataframe, as there are 3 Denny's and 2
LaQuintas, so there are 6 total combinations of the two. The Denny's are 
represented with .x variable names whereas the LaQuintas are represented with 
the .y variable names.

```{r}
haversine <- function(long1, lat1, long2, lat2, round = 3) {
  # convert to radians
  long1 = long1 * pi / 180
  lat1  = lat1  * pi / 180
  long2 = long2 * pi / 180
  lat2  = lat2  * pi / 180
  
  R = 6371 # Earth mean radius in km
  
  a = sin((lat2 - lat1)/2)^2 + cos(lat1) * cos(lat2) * sin((long2 - long1)/2)^2
  d = R * 2 * asin(sqrt(a))
  
  return( round(d,round) ) # distance in km
}
```

```{r}
dn_lq_ak <- dn_lq_ak %>% mutate(distance = haversine(longitude.x, latitude.x, 
                                                     longitude.y, latitude.y))
```

```{r}
dn_lq_ak_mindist <- dn_lq_ak %>%
  group_by(address.x) %>%
  summarise(closest = min(distance))
```

```{r}
summary(dn_lq_ak_mindist)
dn_lq_ak_mindist %>% ggplot(aes(x=closest)) + geom_histogram()
```
There are only 3 data points, so it is difficult to describe the distribution.
All of the Denny's have a LaQuinta within 6km.

##North Carolina 
```{r}
dn_nc <- dennys %>%
  filter(state == "NC")
nrow(dn_nc)
```
There are 28 Denny's in North Carolina.

```{r}
lq_nc <- laquinta %>%
  filter(state == "NC")
nrow(lq_nc)
```
There are 12 LaQuintas in North Carolina.

```{r}
dn_lq_nc <- full_join(dn_nc, lq_nc, by = "state")
dn_lq_nc
```
There are 336 observations in this joined dataframe, as there are 28 Denny's and 
12 LaQuintas. The Denny's are represented with .x variable names whereas the 
LaQuintas are represented with the .y variable names.


```{r}
dn_lq_nc <- dn_lq_nc %>% mutate(distance = haversine(longitude.x, latitude.x, 
                                                     longitude.y, latitude.y))
```

```{r}
dn_lq_nc_mindist <- dn_lq_nc %>%
  group_by(address.x) %>%
  summarise(closest = min(distance))
```

```{r}
summary(dn_lq_nc_mindist)
dn_lq_nc_mindist %>% ggplot(aes(x=closest)) + geom_histogram()
```
The distribution for the closest LaQuinta to each Denny's is skewed right.

##Texas
```{r}
dn_tx <- dennys %>%
  filter(state == "TX")
nrow(dn_tx)
```
There are 200 Denny's in Texas.

```{r}
lq_tx <- laquinta %>%
  filter(state == "TX")
nrow(lq_tx)
```
There are 237 LaQuintas in Texas.

```{r}
dn_lq_tx <- full_join(dn_tx, lq_tx, by = "state")
dn_lq_tx
```
There are 47,400 observations in this joined dataframe, as there are 200 Denny's
and 237 LaQuintas. The Denny's are represented with .x variable names whereas
the LaQuintas are represented with the .y variable names.


```{r}
dn_lq_tx <- dn_lq_tx %>% mutate(distance = haversine(longitude.x, latitude.x, 
                                                     longitude.y, latitude.y))
```

```{r}
dn_lq_tx_mindist <- dn_lq_tx %>%
  group_by(address.x) %>%
  summarise(closest = min(distance))
```

```{r}
summary(dn_lq_tx_mindist)
dn_lq_tx_mindist %>% ggplot(aes(x=closest)) + geom_histogram()
```
The distribution for the closest LaQuinta to each Denny's is skewed right, with
a majority of the Denny's having a LaQuinta within 10 kilometers.

##California
```{r}
dn_ca <- dennys %>%
  filter(state == "CA")
nrow(dn_ca)
```
There are 403 Denny's in California.

```{r}
lq_ca <- laquinta %>%
  filter(state == "CA")
nrow(lq_ca)
```
There are 56 LaQuintas in California.

```{r}
dn_lq_ca <- full_join(dn_ca, lq_ca, by = "state")
dn_lq_ca
```
There are 22,568 observations in this joined dataframe, as there are 403 Denny's
and 56 LaQuintas. The Denny's are represented with .x variable names whereas
the LaQuintas are represented with the .y variable names.


```{r}
dn_lq_ca <- dn_lq_ca %>% mutate(distance = haversine(longitude.x, latitude.x, 
                                                     longitude.y, latitude.y))
```

```{r}
dn_lq_ca_mindist <- dn_lq_ca %>%
  group_by(address.x) %>%
  summarise(closest = min(distance))
```

```{r}
summary(dn_lq_ca_mindist)
dn_lq_ca_mindist %>% ggplot(aes(x=closest)) + geom_histogram()
```
The distribution for the closest LaQuinta to each Denny's is skewed right, with
a majority of the Denny's having a LaQuinta within 25 kilometers.

Among the states we have examined, Mitch Hedberg's joke most applies to Texas,
as most of the observations have less than 10km in between. Alaska has the 
closest, however there are only 3 Denny's, so it is harder to say if that is 
significant, or just a byproduct of Alaska not having many Denny's restaurants.

```{r}
file = here("data_raw","FastFoodRestaurants.csv")
ffr <- read_csv(file)
```

In order to determine if Mitch Hedberg's joke applies to just Denny's, or if it
is just a byproduct of businesses being located in populated areas, we will 
compare the distance between LaQuinta and Denny's to the distance between 
LaQuinta and another fast food restaurant with a similar number of locations.
Because most fast food restaurants have more locations than Denny's does in 
many of the states we are looking at, we will perform a random sample of these
fast food restaurants in order to have a fair comparison.

```{r}
mcd_nc <- ffr %>% 
  rename(state = province) %>%
  filter(state == "NC", name == "McDonald's") %>%
  slice_sample(n=28)

```

```{r}
mcd_lq_nc <- full_join(mcd_nc, lq_nc, by = "state")
mcd_lq_nc
```
There are 336 observations in this joined dataframe, as there are 28 Denny's and 
12 LaQuintas. The Denny's are represented with .x variable names whereas the 
LaQuintas are represented with the .y variable names.


```{r}
mcd_lq_nc <- mcd_lq_nc %>% mutate(distance = haversine(longitude.x, latitude.x, 
                                                     longitude.y, latitude.y))
```

```{r}
mcd_lq_nc_mindist <- mcd_lq_nc %>%
  group_by(address.x) %>%
  summarise(closest = min(distance))
```

```{r}
summary(mcd_lq_nc_mindist)
mcd_lq_nc_mindist %>% ggplot(aes(x=closest)) + geom_histogram()

summary(dn_lq_nc_mindist)
dn_lq_nc_mindist %>% ggplot(aes(x=closest)) + geom_histogram()

```
