---
title: "Data Science in a Box Lab 4"
author: "Alex Lindquist"
date: "10/16/2022"
output: html_notebook
---

##Load Packages

```{r}
library(tidyverse)
library(dsbox)
```

```{r}
file = here("data_raw","states.csv")
states <- read_csv(file)
```

```{r}
dfdenny <- dennys
?dennys
nrow(dennys)
ncol(dennys)
```
1. There are 1643 rows each representing one Denny's restaurant location in the US. 
There are 6 columns each representing a different attribute. The attributes are
address, city, state, zip code, longitude and latitude of the restaurant.

```{r}
dflq <- laquinta
?laquinta
nrow(laquinta)
ncol(laquinta)
```
2. There are 909 rows each representing one LaQuinta motel location in the US. 
There are 6 columns each representing a different attribute. The attributes are
address, city, state, zip code, longitude and latitude of the motel.

3. There are several LaQuinta locations outside of the US, including in Canada,
Mexico, China, New Zealand, Turkey, Chile, Columbia, and the UAE. On the Denny's
website, there are no locations listed outside of the US.

4. In order to determine if there are any locations outside of the US, you could
take the data and filter based on longitude and latitude combinations that are
inside of the US. You could filter based on state abbreviations that are from 
the US, however it is possible that a city abbreviation from another country 
could be the same as a state abbreviation from the US. However it is unlikely,
and it may be easier to just use the city abbreviation.

```{r}
dennys %>%
  filter(!(state %in% states$abbreviation))

dfdenny %>%
  mutate(country = "United States")
```
There are no Denny's outside of the US in this data set.

```{r}
laquinta %>%
  filter(!(state %in% states$abbreviation))

dflq <- dflq %>% mutate(country = case_when(
    state %in% state.abb ~ "United States",
    state %in% c("ON", "BC") ~ "Canada",
    state == "ANT" ~ "Colombia",
    state %in% c("NL", "AG", "QR", "CH", "VE", "PU", "SL") ~ "Mexico",
    state == "FM" ~ "Honduras"))

dflq <- dflq %>%
  filter(country == "United States")
```

```{r}
dfdenny %>%
  count(state) %>%
  inner_join(states, by = c("state" = "abbreviation")) %>%
  mutate(per_square_mile = n / area) %>% 
  arrange(-per_square_mile)
```
DC has the most Denny's per square mile.

```{r}
dflq %>%
  count(state) %>%
  inner_join(states, by = c("state" = "abbreviation")) %>%
  mutate(per_square_mile = n / area) %>% 
  arrange(-per_square_mile)
```
Rhode Island has the most LaQuintas per square mile.

```{r}
dfdenny <- dfdenny %>%
  mutate(establishment = "Denny's")
dflq <- dflq %>%
  mutate(establishment = "La Quinta")

dn_lq <- bind_rows(dfdenny, dflq)

ggplot(dn_lq, mapping = aes(x = longitude, y = latitude, color = establishment)) +
  geom_point()
```

```{r}
dn_lq %>%
  filter(state == "NC") %>% 
  ggplot(dn_lq, mapping = aes(x = longitude, y = latitude, color = establishment)) +
  geom_point(alpha = 0.5)
```

```{r}
dn_lq %>%
  filter(state == "TX") %>% 
  ggplot(dn_lq, mapping = aes(x = longitude, y = latitude, color = establishment)) +
  geom_point(alpha = 0.5)
```


