---
## title: "Airplanes"
## author: "Wynter Malone"
## date: "2022-09-08"
## output: html_notebook

##load packages

```{r}
library(nycflights13)
library(tidyverse)
library(here)
library(janitor)
```

##load flight data

```{r}
df1 = flights
```

##classwork

```{r}
df6 = df1
  filter(dest=="MIA")

df6 %>% ggplot(aes(month)) +geom_bar()
```
  
## Analyzing the graph above, we note:
  
```{r}
df7 = df1
  filter(dest %in% c("LAX", "SAN", "SFO", "ATL", "ORD", "MIA"))

df7 %>% ggplot(aes(month, fill(dest))) +geom_bar()
```


```{r}
thisfile=here("data_raw","DL_SelectFields.zip")

df2019=read_csv(thisfile) %>% clean_names()
```

```{r}

df15 = df2019 %>% filter(passengers>0, origin == "LGA" | dest== "LGA") %>%
  group_by(month, carrier_name)



df15 %>% ggplot(aes(x=month, y=average_pass-per_flight, color=carrier_name))+geom_line()


```

##Some of the airlines don't fly certain months
##One of the pink oneshas the most pass per flight in LGA, might have larger planes.
##Plane companies on the lower half of the graph might be smaller planes/ private jets
##most companies average around 40-60 or 100-120 passengers per flight
##more passengers in the summer and Nov + Dec




